---
title: "IE 582"
author: "Burak Çetiner"
date: "2023-11-01"
output: html_document
---

---
title: "IE 582 Statistical Learning for Data Mining"
author: "Burak Çetiner"
date: "2023-11-01"
output: html_document
---

## Homework 1

```{r setup, include=FALSE}
library(corrr)
library(ggcorrplot)
library(FactoMineR)
library(devtools)
library(ggplot2)
library(factoextra)
library(dplyr)
library(readxl)
library(httr)
```



```{r}
github_url <- "https://github.com/BU-IE-582/fall-23-burakcetiner3/raw/main/all_ticks_wide.xlsx"
local_file_path <- "all_ticks_wide.xlsx"
download.file(url = github_url, destfile = local_file_path, mode = "wb")
Dataset <- read_excel(local_file_path)
summary(Dataset)
```
First I wanted to see the Five Number Summary Statistics (summary) of each dataset. I realized that there are many missing values (NAs) and I decide to get rid off them


```{r}
Dataset1 = Dataset[complete.cases(Dataset),] #removing NAs
```

Then instead of dealing with 61 different stocks, I decide to choose 10 of them. I chose the foloowing 10 stocks because I have an insight about them due to my brother's trades and advices on Turkish financial market.

```{r}
AEFES = Dataset1$AEFES
AKBNK = Dataset1$AKBNK
AKSA = Dataset1$AKSA
AKSEN = Dataset1$AKSEN
SASA = Dataset1$SASA
CCOLA = Dataset1$CCOLA
GARAN = Dataset1$GARAN
MGROS = Dataset1$MGROS 
TCELL = Dataset1$TCELL
TUPRS = Dataset1$TUPRS
```


## 4.1 Descriptive Analysis

First I will combine them to see the correlation relationships

```{r}
Table = cbind(AKBNK,AKSA,AKSEN,SASA,CCOLA,GARAN,MGROS,TCELL, TUPRS) # I construct a new table to see the pairs
pairs(Table) # scatter diagrams of each pairs
```

It seems like there is positive strong correlation between AKBNK and GARAN. Let's investigate further

```{r}
plot(AKBNK, GARAN)
round(cor(AKBNK, GARAN),3) # r = 0.907 indicates a strong positive correlation between AKBNK and GARAN.
summary(AKBNK)
summary(GARAN)
boxplot(AKBNK) # This function visualize 5-number-summary statistics of AKBNK stock
boxplot(GARAN) # This function visualize 5-number-summary statistics of GARAN stock
```

To finalize descriptive statistics of each stock we have to define standard deviation, variance, mode, interquartile range, and finally the histogram of each stock.

```{r}
mode_AKBNK = sort(table(AKBNK), decreasing=T)
mode_AKBNK[1] #mode of AKBNK stock is 6.79
sd(AKBNK) #standard deviation of AKBNK
var(AKBNK) #variance of AKBNK
IQR(AKBNK) #interquartile range of AKBNK
hist(AKBNK, main="The histogram of stock AKBNK", col="skyblue")
```



```{r}
mode_GARAN = sort(table(GARAN), decreasing=T)
mode_GARAN[1] #mode of GARAN stock is 8.95
sd(GARAN) #standard deviation of GARAN
var(GARAN) #variance of GARAN
IQR(GARAN) #interquartile range of GARAN
hist(GARAN,main="The histogram of stock GARAN", col="skyblue")
```

2nd finding:
It seems like there is positive correlation between AKSA and AKSEN. Let's investigate further

```{r}
summary(AKSA)
summary(AKSEN)
boxplot(AKSA) # This function visualize 5-number-summary statistics of AKSA stock
boxplot(AKSEN) # This function visualize 5-number-summary statistics of AKSEN stock
plot(AKSA, AKSEN)
round(cor(AKSA, AKSEN),3) # 0.636 indicates a positive correlation between AKSA and AKSEN.
```

To finalize descriptive statistics of each stock we have to define standard deviation, variance, mode, interquartile range, and finally the histogram of each stock.

```{r}
mode_AKSA = sort(table(AKSA), decreasing=T)
mode_AKSA[1] #mode of AKSA stock is 8.0397
sd(AKSA) #standard deviation of AKSA
var(AKSA) #variance of AKSA
IQR(AKSA) #interquartile range of AKSA
hist(AKSA, main="The histogram of stock AKSA", col="skyblue")
```

```{r}
mode_AKSEN = sort(table(AKSEN), decreasing=T)
mode_AKSEN[1] #mode of AKSEN stock is 2.35
sd(AKSEN) #standard deviation of AKSEN
var(AKSEN) #variance of AKSEN
IQR(AKSEN) #interquartile range of AKSEN
hist(AKSEN, main="The histogram of stock AKSEN", col="skyblue")
```

## 4.2 Moving Window Correlation

```{r}
AKBNK_april = AKBNK[11:58]; AKBNK_april
AKBNK_may = AKBNK[59:101]
AKBNK_june = AKBNK[102:150]
AKBNK_july = AKBNK[151:182]
AKBNK_august = AKBNK[183:224]
AKBNK_september = AKBNK[225:269]

GARAN_april = GARAN[11:58]; GARAN_april
GARAN_may = GARAN[59:101]
GARAN_june = GARAN[102:150]
GARAN_july = GARAN[151:182]
GARAN_august = GARAN[183:224]
GARAN_september = GARAN[225:269]

plot(GARAN_april, AKBNK_april)
cor(GARAN_april, AKBNK_april)

plot(GARAN_may, AKBNK_may)
cor(GARAN_may, AKBNK_may)

plot(GARAN_june, AKBNK_june)
cor(GARAN_june, AKBNK_june)

plot(GARAN_july, AKBNK_july)
cor(GARAN_july, AKBNK_july)

plot(GARAN_august, AKBNK_august)
cor(GARAN_august, AKBNK_august)

plot(GARAN_september, AKBNK_september)
cor(GARAN_september, AKBNK_september)
```

```{r}
AKBNK_2016 = AKBNK[1:376]; AKBNK_2016
AKBNK_2017 = AKBNK[377:1055]
AKBNK_2018 = AKBNK[1056:5318]
AKBNK_2019 = AKBNK[5319:9228]

GARAN_2016 = GARAN[1:376]; GARAN_2016
GARAN_2017 = GARAN[377:1055]
GARAN_2018 = GARAN[1056:5318]
GARAN_2019 = GARAN[5319:9228]

plot(GARAN_2016, AKBNK_2016)
cor(GARAN_2016, AKBNK_2016)

plot(GARAN_2017, AKBNK_2017)
cor(GARAN_2017, AKBNK_2017)

plot(GARAN_2018, AKBNK_2018)
cor(GARAN_2018, AKBNK_2018)

plot(GARAN_2019, AKBNK_2019)
cor(GARAN_2019, AKBNK_2019)
```

# 4.3 Principal Component Analysis (PCA)

```{r}
numerical_data = Dataset1[2:61]
data_normalized <- scale(numerical_data)
head(data_normalized)
```

First I created the correlation matrix by using thw whole dataset.

```{r}
corr_matrix <- cor(data_normalized)
ggcorrplot(corr_matrix)

```

However, we cannot comment much about it. It does not help much, since there are 61 variables. So, I decide to continue principal compenent analysis by using 10 preffered stocks.

```{r}
Preffered_stock = Table
Preffered_normalized <- scale(Table)
```

```{r}
corr_matrix2 <- cor(Preffered_normalized) ;corr_matrix2
ggcorrplot(corr_matrix2)

data.pca <- princomp(corr_matrix2)
summary(data.pca)
```

We see that highest correlation value appers in between AKBNK and GARAN as we earlier found out.


```{r}
data.pca$loadings[, 1:2]
fviz_screeplot(data.pca, addlabels = TRUE)
```


```{r}
fviz_pca_var(data.pca, col.var = "black")
fviz_cos2(data.pca, choice = "var", axes = 1:2)
fviz_pca_var(data.pca, col.var = "cos2",
             gradient.cols = c("black", "orange", "green"),
             repel = TRUE)
```


## 4.4 Inference with Google Trends

You can find the related figures and the following comment on Google Trend Analysis in the report file.
